Function to find association b/w n significant eigen genes of meta modules with Diagnosis, Diagnosis.Gender, ApoE

### 2017-10-10 code/Rmd/networkAssociationTest_metaModules.Rmd
Temp commit

EXPR = downloadFile('syn10160584')
all.used.ids = c(all.used.ids, 'syn10160584')
gene.id.map = data.frame(ID = EXPR$Gene.ID, Gene.ID = EXPR$Gene.ID) %>%
  tidyr::separate(ID, c('GeneID','position'), sep = '\\.')
EXPR = left_join(EXPR, gene.id.map)
rownames(EXPR) = EXPR$GeneID
EXPR$Gene.ID = NULL
EXPR$GeneID = NULL

# MODULE.OVRLP = synTableQuery('SELECT * FROM syn10339153')@values
# all.used.ids = c(all.used.ids, 'syn10339153')
# MODULE.ENRICH = synTableQuery('SELECT * FROM syn10492048')@values
# all.used.ids = c(all.used.ids, 'syn10492048')
# 
# MAGMA.ENRICH = synTableQuery('SELECT * FROM syn10380432')@values
# all.used.ids = c(all.used.ids, 'syn10380432')
```

@@

    Heatmap(mdl.ovrlp, col = colorRamp2(seq(0,1,length.out = 5), viridis::viridis(5)),
            name = 'JacDist', cell_fun = function(j,i,x,y,wifht,height,fill){
              grid.text(mdl.ovrlp.log[i,j] %>% format(digits = 1), x, y)
            })
  }
```

### Associaiton with Diagnosis
```{r Dx, fig.height=5, fig.width=10}
# Find assocition b/w Diagnosis category with n eigen genes
FORMULA = as.formula('PC~Diagnosis+Batch+RIN+PCT_INTRONIC_BASES+PMI+PCT_INTERGENIC_BASES+AOD+PCT_PF_READS_ALIGNED+Gender')
assoc.stat.diagnosis = MODULES %>%
  dplyr::filter(GeneID %in% rownames(EXPR)) %>%
  plyr::dlply(.(ModuleName, brainRegion), .fun = findAssociationStatistics, 
              COVARIATES, EXPR, FORMULA,
              .parallel = T, .paropts = list(.packages = c('dplyr', 'data.table')))

@@

writeLines('Percent of modules whose PC1 is significantly associated')
Dx %>%
  dplyr::group_by(brainRegion) %>%
  dplyr::summarise(nsig = round(sum(pc1.fdr <= 0.05)/sum(pc1.fdr <= 1) * 100, 2)) %>%
  kable()
```

### Associaiton with Diagnosis.Gender
```{r Dx.Gender, fig.height=5, fig.width=10}
# Find assocition b/w Diagnosis.Gender category with n eigen genes
COVARIATES$Diagnosis.Gender = paste(COVARIATES$Diagnosis, COVARIATES$Gender, sep= '.') %>%
  as.factor
FORMULA = as.formula('PC~Diagnosis.Gender+Batch+RIN+PCT_INTRONIC_BASES+PMI+PCT_INTERGENIC_BASES+AOD+PCT_PF_READS_ALIGNED')
assoc.stat.diagnosis.gender = MODULES %>%
  dplyr::filter(GeneID %in% rownames(EXPR)) %>%
  plyr::dlply(.(ModuleName, brainRegion), .fun = findAssociationStatistics, 
              COVARIATES, EXPR, FORMULA,
              .parallel = T, .paropts = list(.packages = c('dplyr', 'data.table')))

@@

writeLines('Percent of modules whose PC1 is significantly associated')
Dx.gender %>%
  dplyr::group_by(brainRegion) %>%
  dplyr::summarise(nsig = round(sum(pc1.fdr <= 0.05)/sum(pc1.fdr <= 1) * 100, 2)) %>%
  kable()
```

### Enrichment for differential expression
```{r Dx.enrich}
# Get differential expression gene list from synapse
load(synGet('syn10496554')@filePath)
all.used.ids = c(all.used.ids, 'syn10496554')
names(amp.ad.de.geneSets) = gsub('CER','CBE', names(amp.ad.de.geneSets))

@@

  enrichResults = plyr::ldply(modMembers, .fun = function(y, deGenes, backgroundGenes){
    plyr::ldply(deGenes, .fun = function(z, y, backgroundGenes){
      fisherEnrichment(y,z,backgroundGenes)
    }, y, backgroundGenes, .id = 'Dx.GeneSet')
  }, deGenes, backgroundGenes, .id = 'ModuleNameFull') %>%
    dplyr::mutate(fdr = p.adjust(pval))
}, .id = 'brainRegion')
```
### Enrichment for gene sets
```{r pathway.enrich}
# Download all related genesets from synapse
load(synGet('syn4867851')@filePath)
all.used.ids = c(all.used.ids, 'syn4867851')

@@

  enrichResults = plyr::ldply(geneSetRef, .fun = function(y, modMembers, backgroundGenes){
    fisherEnrichment(y, modMembers, backgroundGenes)
  }, modMembers, backgroundGenes, .id = 'GeneSetName', .parallel = T, 
  .paropts = list(.packages = c('CovariateAnalysis'))) %>%
    dplyr::mutate(fdr = p.adjust(pval))
}, 
.id = 'ModuleNameFull', geneSet.ref, backgroundGenes)
```

### Find overlap between modules in different brainregion
```{r module.overlap}

@@

### Intereating modules
```{r int.modules}

### Store results in synapse
```{r syn.store, include=FALSE, eval = F}
write.table(Dx, file = 'AssociaitonStatWithDx.tsv', sep = '\t', quote = F, row.names = F)
OBJ = File('AssociaitonStatWithDx.tsv', name = 'Association with Dx', parentId = 'syn11140497')
OBJ = synStore(OBJ, used = all.used.ids, executed = thisFile, activityName = activityName)

>>>>

EXPR = downloadFile('syn10160584')
all.used.ids = c(all.used.ids, 'syn10160584')
gene.id.map = data.frame(ID = EXPR$Gene.ID, Gene.ID = EXPR$Gene.ID) %>%
  tidyr::separate(ID, c('GeneID','position'), sep = '\\.')
EXPR = left_join(EXPR, gene.id.map)
rownames(EXPR) = EXPR$GeneID
EXPR$Gene.ID = NULL
EXPR$GeneID = NULL
```

@@

    Heatmap(mdl.ovrlp, col = colorRamp2(seq(0,1,length.out = 5), viridis::viridis(5)),
            name = 'JacDist', cell_fun = function(j,i,x,y,wifht,height,fill){
              grid.text(mdl.ovrlp.log[i,j] %>% format(digits = 1), x, y)
            })
  }
```
### Associaiton with Diagnosis
```{r Dx, fig.height=5, fig.width=10}
# Find assocition b/w Diagnosis category with n eigen genes
FORMULA = as.formula('PC~Diagnosis+Batch+RIN+PCT_INTRONIC_BASES+PMI+PCT_INTERGENIC_BASES+AOD+PCT_PF_READS_ALIGNED+Gender')
assoc.stat.diagnosis = MODULES %>%
  dplyr::filter(GeneID %in% rownames(EXPR)) %>%
  plyr::dlply(.(ModuleName, brainRegion), .fun = findAssociationStatistics, 
              COVARIATES, EXPR, FORMULA,
              .parallel = T, .paropts = list(.packages = c('dplyr', 'data.table')))

@@

writeLines('Percent of modules whose PC1 is significantly associated')
Dx %>%
  dplyr::group_by(brainRegion) %>%
  dplyr::summarise(nsig = round(sum(pc1.fdr <= 0.05)/sum(pc1.fdr <= 1) * 100, 2)) %>%
  kable()
```
### Associaiton with Diagnosis.Gender
```{r Dx.Gender, fig.height=10, fig.width=10}
# Find assocition b/w Diagnosis.Gender category with n eigen genes
COVARIATES$Diagnosis.Gender = paste(COVARIATES$Diagnosis, COVARIATES$Gender, sep= '.') %>%
  as.factor
FORMULA = as.formula('PC~Diagnosis.Gender+Batch+RIN+PCT_INTRONIC_BASES+PMI+PCT_INTERGENIC_BASES+AOD+PCT_PF_READS_ALIGNED')
assoc.stat.diagnosis.gender = MODULES %>%
  dplyr::filter(GeneID %in% rownames(EXPR)) %>%
  plyr::dlply(.(ModuleName, brainRegion), .fun = findAssociationStatistics, 
              COVARIATES, EXPR, FORMULA,
              .parallel = T, .paropts = list(.packages = c('dplyr', 'data.table')))

@@

writeLines('Percent of modules whose PC1 is significantly associated')
Dx.gender %>%
  dplyr::group_by(brainRegion) %>%
  dplyr::summarise(nsig = round(sum(pc1.fdr <= 0.05)/sum(pc1.fdr <= 1) * 100, 2)) %>%
  kable()

writeLines('Comparison b/w Dx and Dx.gender')
tmp = dplyr::left_join(Dx %>%
                         dplyr::select(ModuleName, brainRegion, total.eff.sz, total.fdr) %>%
                         dplyr::rename(Dx = total.eff.sz, Dx.fdr = total.fdr),
                       Dx.gender %>%
                         dplyr::select(ModuleName, brainRegion, total.eff.sz, total.fdr) %>%
                         dplyr::rename(Dx.gender = total.eff.sz, Dx.gender.fdr = total.fdr))
p = ggplot(tmp, aes(x = Dx, y = Dx.gender, color = -log10(Dx.fdr+1e-100))) + geom_point() + facet_grid(.~brainRegion) + geom_smooth(method = 'lm')
p = p +scale_color_continuous(low = 'white', high = 'red')
p
```
### Enrichment for differential expression
```{r Dx.enrich, fig.height=10, fig.width=10}
# Get differential expression gene list from synapse
load(synGet('syn10496554')@filePath)
all.used.ids = c(all.used.ids, 'syn10496554')
names(amp.ad.de.geneSets) = gsub('CER','CBE', names(amp.ad.de.geneSets))

@@

  enrichResults = plyr::ldply(modMembers, .fun = function(y, deGenes, backgroundGenes){
    plyr::ldply(deGenes, .fun = function(z, y, backgroundGenes){
      fisherEnrichment(y,z,backgroundGenes)
    }, y, backgroundGenes, .id = 'Dx.GeneSet')
  }, deGenes, backgroundGenes, .id = 'ModuleNameFull') %>%
    dplyr::mutate(fdr = p.adjust(pval))
}, .id = 'brainRegion')

writeLines('Comparison b/w Dx and Dx.gender')
Dx.enrich1 = Dx.enrich %>%
  dplyr::select(-Genes) %>%
  tidyr::separate(Dx.GeneSet, c('brainRegion', 'Comparison', 'Gender', 'Direction'), sep = '\\.')
Dx.enrich1$Direction[is.na(Dx.enrich1$Direction)] = Dx.enrich1$Gender[is.na(Dx.enrich1$Direction)]
Dx.enrich1$Gender = gsub('UP', 'ALL', Dx.enrich1$Gender)
Dx.enrich1$Gender = gsub('DOWN', 'ALL', Dx.enrich1$Gender)
Dx.enrich1 = dplyr::filter(Dx.enrich1, Comparison == 'AD-CONTROL', fdr <= 0.05)
p = ggplot(Dx.enrich1, aes(x = Gender, y = -log10(fdr))) + geom_boxplot() 
p = p + coord_flip() + facet_grid(.~brainRegion, scales = 'free') + geom_jitter(aes(color = Direction))
pl = list();pl[[1]] = p

p = ggplot(Dx.enrich1, aes(x = Gender, y = Odds.Ratio)) + geom_boxplot() 
p = p + coord_flip() + facet_grid(.~brainRegion, scales = 'free') + geom_jitter(aes(color = Direction))
pl[[2]] = p
multiplot(plotlist = pl, cols = 1)
```
### Enrichment for gene sets
```{r pathway.enrich, results='asis'}
# Download all related genesets from synapse
load(synGet('syn4867851')@filePath)
all.used.ids = c(all.used.ids, 'syn4867851')

@@

  enrichResults = plyr::ldply(geneSetRef, .fun = function(y, modMembers, backgroundGenes){
    fisherEnrichment(y, modMembers, backgroundGenes)
  }, modMembers, backgroundGenes, .id = 'GeneSetName', .parallel = T, 
  .paropts = list(.packages = c('CovariateAnalysis'))) %>%
    dplyr::mutate(fdr = p.adjust(pval))
}, 
.id = 'ModuleNameFull', geneSet.ref, backgroundGenes)

# Get AD specific pathways
other.ad.pathways = c("KEGG_2016.Alzheimer's disease_Homo sapiens_hsa05010",
                      "WikiPathways_2016.Alzheimers Disease_Homo sapiens_WP2059")

writeLines('AD related pathways enriched an fdr of 0.05')
ad.enrich = Pathway.enrich %>%
  dplyr::select(ModuleNameFull, GeneSetName, Odds.Ratio, fdr) %>%
  dplyr::filter(GeneSetName %in% other.ad.pathways, fdr <= 0.05) %>%
  dplyr::left_join(MODULES %>%
                     dplyr::select(ModuleNameFull, brainRegion)) %>%
  unique()
kable(ad.enrich)
```
### Find overlap between modules in different brainregion
```{r module.overlap, fig.height=5, fig.width=10}
# FisherEnrichment b/w modules
MOD.MEMBERS = plyr::dlply(MODULES, .(ModuleNameFull), .fun = function(x) unique(x$external_gene_name))
MOD.OVRLP = plyr::ldply(MOD.MEMBERS, .fun = function(x, modMembers, backgroundGenes){
  plyr::ldply(modMembers, .fun = function(y, x, backgroundGenes){
    fisherEnrichment(y, x, backgroundGenes)
  }, x, backgroundGenes, .parallel = T, .paropts = list(.packages = c('CovariateAnalysis')), .id = 'MOD1') %>%
    dplyr::rename(MOD1 = ModuleNameFull)
}, MOD.MEMBERS, backgroundGenes, .id = 'MOD2') %>%
  dplyr::select(-Genes) %>%
  dplyr::rename(MOD2 = ModuleNameFull)
```
### Interesting modules
```{r int.modules, fig.height=7, fig.width=7, results='asis'}
# Wrt case control
tmp = inner_join(Dx %>% 
                   dplyr::filter(total.fdr <= 0.05) %>%
                   dplyr::select(ModuleNameFull, brainRegion, total.eff.sz, total.fdr) %>%
                   dplyr::rename(assoc.eff.sz = total.eff.sz, assoc.fdr = total.fdr),
                 Dx.enrich1 %>%
                   dplyr::filter(fdr <= 0.05, Gender == 'ALL') %>%
                   dplyr::select(ModuleNameFull, brainRegion, Direction, Odds.Ratio, fdr) %>%
                   dplyr::rename(diffexp.fdr = fdr, diffexp.or = Odds.Ratio)) %>%
  left_join(ad.enrich %>%
              dplyr::select(ModuleNameFull, fdr, brainRegion) %>%
              dplyr::rename(ad.pathway.fdr = fdr))
writeLines('Imporatant Modules for case control comparison')
kable(tmp)

writeLines('Up regulated modules')
tmp1 = filter(tmp, Direction == 'UP')
tmp1 = filter(MOD.OVRLP, MOD1 %in% tmp1$ModuleNameFull, MOD2 %in% tmp1$ModuleNameFull)
g = igraph::graph_from_data_frame(tmp1)
tmp1.or = igraph::as_adj(g, attr = 'Odds.Ratio') %>% data.matrix
tmp1.or[is.infinite(tmp1.or)] = 0

tmp1.fdr = igraph::as_adj(g, attr = 'pval') %>% data.matrix()
tmp1.fdr = tmp1.fdr <= 0.05
tmp1.fdr[tmp1.fdr] = '*'
tmp1.fdr[tmp1.fdr == 'FALSE'] = ''

h = ComplexHeatmap::Heatmap(log2(tmp1.or+1e-2), col = viridis::viridis(5), 
                        cell_fun = function(j,i,x,y,wifht,height,fill){
                          grid.text(tmp1.fdr[i,j] %>% format(digits = 1), x, y)
                        }, name = 'log2(OR)')
draw(h, heatmap_legend_side = 'left')

dplyr::filter(Pathway.enrich, 
              GeneSetName %in% grep('CellType', unique(Pathway.enrich$GeneSetName), value = T),
              fdr <= 0.05, 
              ModuleNameFull %in% rownames(tmp1.or)) %>%
  dplyr::select(-Genes)

writeLines('Down regulated modules')
tmp1 = filter(tmp, Direction == 'DOWN')
tmp1 = filter(MOD.OVRLP, MOD1 %in% tmp1$ModuleNameFull, MOD2 %in% tmp1$ModuleNameFull)
g = igraph::graph_from_data_frame(tmp1)
tmp1.or = igraph::as_adj(g, attr = 'Odds.Ratio') %>% data.matrix
tmp1.or[is.infinite(tmp1.or)] = 0

tmp1.fdr = igraph::as_adj(g, attr = 'pval') %>% data.matrix()
tmp1.fdr = tmp1.fdr <= 0.05
tmp1.fdr[tmp1.fdr] = '*'
tmp1.fdr[tmp1.fdr == 'FALSE'] = ''

h = ComplexHeatmap::Heatmap(log2(tmp1.or+1e-2), col = viridis::viridis(5), 
                        cell_fun = function(j,i,x,y,wifht,height,fill){
                          grid.text(tmp1.fdr[i,j] %>% format(digits = 1), x, y)
                        }, name = 'log2(OR)')
draw(h, heatmap_legend_side = 'left')

dplyr::filter(Pathway.enrich, 
              GeneSetName %in% grep('CellType', unique(Pathway.enrich$GeneSetName), value = T),
              fdr <= 0.05, 
              ModuleNameFull %in% rownames(tmp1.or)) %>%
  dplyr::select(-Genes)

@@

### Store results in synapse
```{r syn.store, include=FALSE, eval = F, cache=FALSE}
write.table(Dx, file = 'AssociaitonStatWithDx.tsv', sep = '\t', quote = F, row.names = F)
OBJ = File('AssociaitonStatWithDx.tsv', name = 'Association with Dx', parentId = 'syn11140497')
OBJ = synStore(OBJ, used = all.used.ids, executed = thisFile, activityName = activityName)
